|

### 2023-12-07 chapter_2/src/manuscript.qmd
doing a bit more drafting

Do colors actually shape political attitudes among the mass public? There are two goals that I am hoping to achieve in this chapter. The first is to convince the reader that the colors red and blue convey consistent, affiliative information. The second is that these connections between colors and political groups can indeed affect political outcomes.

Many understand -- in a general sense -- that "colors matter" when it comes to attitudes and behaviors. Color theorists and scholars of marketing understand color as having important implications for reactions to a product and resulting consumer behavior. For example, @kuo_impact_2023 demonstrate that hotel rooms that use more "cool colors" (blues, greens, etc.) evoke positive affect for potential customers and that these positive affective responses are predictive of their eventual choice to book the room. Colors mean a varying number of things, however.

@@

In this chapter, I build upon descriptive evidence suggesting that Democratic campaigns use the color blue more than Republican campaigns [@williams_whats_2022] and that Republican voters express a preference for red during election season [@schloss_politics_2014]. I present a snap-judgment model of politically-relevant visual information processing that outlines a) the cognitive processes that drive these associations and b) notes how those associations drive important political outcomes. I draw upon theories of associative memory and spreading activation [see @collins_spreading-activation_1975] to conceptualize the association of the colors red and blue with Republicans and Democrats, thinking of them as as two interconnected nodes. I conceptualize these connections as carrying affective information -- an idea informed by perspectives on biological survival reaching back to Darwin [@ralph_neuroscience_2018]. When these nodes are activated by exposure to these colors in political contexts, other related nodes become active and contain affective information. As partisanship evokes strong affective reactions for many members of the public [@iyengar_affect_2012], this activation elicits an affective reaction. Considering that most visual information -- including these processes of association -- occur pre-cognitively [@ames_impression_2012;@fazio_attitudes_2007;@sander_cambridge_2013], I argue that the elicited affective response to such information can be "transferred" [see @morris_activation_2003;@taber_illusion_2016] to later post-cognitive processing of more traditional forms of political information such as stated policy positions. The implication of this is that these associations can *pre-cognitively* inform reactions to political information and political behavior

>>>>

Do colors actually shape political attitudes among the mass public? Using the model covered in [Chapter 1](../../chapter_1/src/manuscript.qmd), there are two goals that I am hoping to achieve in this chapter. The first is to convince the reader that the colors red and blue convey consistent, affiliative information. The second is that these connections between colors and political groups can indeed affect political outcomes.

Many understand -- in a general sense -- that "colors matter" when it comes to attitudes and behaviors. As discussed in the previous chapter, color theorists and scholars of marketing understand color as having important implications for reactions to a product and resulting consumer behavior. For example, @kuo_impact_2023 demonstrate that hotel rooms that use more "cool colors" (blues, greens, etc.) evoke positive affect for potential customers and that these positive affective responses are predictive of their eventual choice to book the room. Colors mean a varying number of things, however.

@@

In this chapter, I build upon descriptive evidence suggesting that Democratic campaigns use the color blue more than Republican campaigns [@williams_whats_2022] and that Republican voters express a preference for red during election season [@schloss_politics_2014]. I present a first test of the snap-judgment model of politically-relevant visual information processing that outlines a) the cognitive processes that drive these associations and b) notes how those associations drive important political outcomes. As discussed in [Chapter 1](../../chapter_1/src/manuscript.qmd), I draw upon theories of associative memory and spreading activation [see @collins_spreading-activation_1975] to conceptualize the association of the colors red and blue with Republicans and Democrats, thinking of them as as two interconnected nodes. I conceptualize these connections as carrying affective information -- an idea informed by perspectives on biological survival reaching back to Darwin [@ralph_neuroscience_2018]. When these nodes are activated by exposure to these colors in political contexts, other related nodes become active and contain affective information. As partisanship evokes strong affective reactions for many members of the public [@iyengar_affect_2012], this activation elicits an affective reaction. Considering that most visual information -- including these processes of association -- occur pre-cognitively [@ames_impression_2012;@fazio_attitudes_2007;@sander_cambridge_2013], I argue that the elicited affective response to such information can be "transferred" [see @morris_activation_2003;@taber_illusion_2016] to later post-cognitive processing of more traditional forms of political information such as stated policy positions. The implication of this is that these associations can *pre-cognitively* inform reactions to political information and political behavior
